---
title: "p8105_hw6_jz3902"
author: "Jinghan Zhao"
date: "2024-11-23"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%",
  warning = FALSE,
  message = FALSE
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Importing data.

```{r import_weather}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Writing functions to extract $\hat{r}^2$ and generate $log(\hat{\beta_0} * \hat{\beta_1})$.

```{r r_squared}
r_squared = function (input_fit){
  
  r_squared = 
    input_fit %>% 
    broom::glance() %>% 
    pull(r.squared)
  
  return(r_squared)
}
```

```{r log_beta}
log_beta = function (input_fit){
  
  beta_0 = 
    input_fit %>% 
    broom::tidy() %>% 
    filter(term == "(Intercept)") %>% 
    pull(estimate)
  
  beta_1 = 
    input_fit %>% 
    broom::tidy() %>% 
    filter(term == "tmin") %>% 
    pull(estimate)
  
  log_beta0_beta1 = log(beta_0 * beta_1)
  
  return(log_beta0_beta1)
}
```

Doing bootstrapping and produce the two quantities.

```{r bootstrap}
boot_straps = 
  weather_df %>% 
  bootstrap(5000) %>% 
  mutate(
    strap = map(strap, as_tibble),
    models = map(strap, \(df) lm(tmax ~ tmin, data = df)),
    r_squared = map(models, \(fit) r_squared(input_fit = fit)),
    log_beta0_beta1 = map(models, \(fit) log_beta(input_fit = fit))
  ) %>% 
  select(.id, r_squared, log_beta0_beta1) %>% 
  unnest(c(r_squared, log_beta0_beta1))
```

### $\hat{r}^2$: Distribution and CI

```{r plot_ci_r}
boot_straps %>% 
  ggplot(aes(x = r_squared)) +
  geom_density()

boot_straps %>% 
  select(r_squared) %>% 
  t.test(conf.level = 0.95) %>% 
  broom::tidy() %>% 
  select(conf.low, conf.high)
```

The distribution of $\hat{r}^2$ is approximately normal and slightly left-skewed. This skewness may be caused by the randomness of sampling.

### $log(\hat{\beta_0} * \hat{\beta_1})$: Distribution and CI

```{r plot_ci_beta}
boot_straps %>% 
  ggplot(aes(x = log_beta0_beta1)) +
  geom_density()

boot_straps %>% 
  select(log_beta0_beta1) %>% 
  t.test(conf.level = 0.95) %>% 
  broom::tidy() %>% 
  select(conf.low, conf.high)
```

The distribution of $log(\hat{\beta_0} * \hat{\beta_1})$ is approximately normal.

