---
title: "p8105_hw6_jz3902"
author: "Jinghan Zhao"
date: "2024-11-23"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .8,
  out.width = "90%",
  warning = FALSE,
  message = FALSE
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Importing data.

```{r import_weather}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Writing functions to extract $\hat{r}^2$ and generate $log(\hat{\beta_0} * \hat{\beta_1})$.

```{r r_squared}
r_squared = function (input_fit){
  
  r_squared = 
    input_fit %>% 
    broom::glance() %>% 
    pull(r.squared)
  
  return(r_squared)
}
```

```{r log_beta}
log_beta = function (input_fit){
  
  beta_0 = 
    input_fit %>% 
    broom::tidy() %>% 
    filter(term == "(Intercept)") %>% 
    pull(estimate)
  
  beta_1 = 
    input_fit %>% 
    broom::tidy() %>% 
    filter(term == "tmin") %>% 
    pull(estimate)
  
  log_beta0_beta1 = log(beta_0 * beta_1)
  
  return(log_beta0_beta1)
}
```

Doing bootstrapping and produce the two quantities.

```{r bootstrap}
boot_straps = 
  weather_df %>% 
  bootstrap(5000) %>% 
  mutate(
    strap = map(strap, as_tibble),
    models = map(strap, \(df) lm(tmax ~ tmin, data = df)),
    r_squared = map(models, \(fit) r_squared(input_fit = fit)),
    log_beta0_beta1 = map(models, \(fit) log_beta(input_fit = fit))
  ) %>% 
  select(.id, r_squared, log_beta0_beta1) %>% 
  unnest(c(r_squared, log_beta0_beta1))
```

### $\hat{r}^2$: Distribution and CI

```{r plot_ci_r}
boot_straps %>% 
  ggplot(aes(x = r_squared)) +
  geom_density()

boot_straps %>% 
  select(r_squared) %>% 
  t.test(conf.level = 0.95) %>% 
  broom::tidy() %>% 
  select(conf.low, conf.high)
```

The distribution of $\hat{r}^2$ is approximately normal and slightly left-skewed. This skewness may be caused by the randomness of sampling.

### $log(\hat{\beta_0} * \hat{\beta_1})$: Distribution and CI

```{r plot_ci_beta}
boot_straps %>% 
  ggplot(aes(x = log_beta0_beta1)) +
  geom_density()

boot_straps %>% 
  select(log_beta0_beta1) %>% 
  t.test(conf.level = 0.95) %>% 
  broom::tidy() %>% 
  select(conf.low, conf.high)
```

The distribution of $log(\hat{\beta_0} * \hat{\beta_1})$ is approximately normal.


## Problem 2

```{r import_homicide}
homicide_df = 
  read_csv("data/homicide-data.csv") %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age)
    ) %>% 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
  )
```

### Baltimore: Logistic regression

```{r fit_baltimore}
fit_baltimore = 
  homicide_df %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) 

fit_baltimore %>% 
  broom::tidy()
  
fit_baltimore %>% 
  broom::tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  filter(term == "victim_sexMale") %>% 
  select(term, estimate, conf.low, conf.high) %>% 
  rename(odds_ratio = estimate)
```

Use `exponentiate = TRUE` to compute odds ratio.

Homicides in which the victim is male are significantly less likely to be resolved than those in which the victim is female.


### All cities: Logistic regression

```{r fit_cities}
fit_cities = 
  homicide_df %>% 
  select(city_state, resolved, victim_age, victim_sex, victim_race) %>% 
  filter(victim_sex %in% c("Male", "Female")) %>% 
  nest(data = resolved:victim_race) %>% 
  mutate(
    models = map(data, \(df) glm(resolved ~ victim_age + victim_sex + victim_race, 
                                 data = df, family = binomial())),
    results = map(models, \(x) broom::tidy(x = x, conf.int = TRUE, exponentiate = TRUE))
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  rename(odds_ratio = estimate)

fit_cities %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, odds_ratio)) %>% 
  ggplot(aes(x = city_state, y = odds_ratio, color = city_state)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  labs(
    title = "Solving homicides comparing male victims to female victims",
    x = "City",
    y = "Odds Ratio",
    caption = "Estimated odds ratio and CI for solving homicides comparing male victims to female victims.") +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 6, angle = 60, hjust = 1))
```

* For most cities in the dataset, homicides in which the victim is male are less or similarly likely to be resolved than those in which the victim is female.

* For Fresno, CA, Stockton, CA and Albuquerque, NM, the estimated odds ratios are greater than 1, while the confidence intervals still include 1, meaning that we do not have enough evidence to conclude that homicides with male victims are more likely to be resolved than homicides with female victims.
